---
title: "Project 2 Welcome"
author: "Brian Sugg"
date: "7/3/2020"
output:
  rmarkdown::github_document:
    toc: true
    toc_depth: 3
---

```{r setupREADME, include=FALSE}
library(tidyverse)
library(Lahman)
library(formatR)
library(dplyr)
library(knitr)
library(class)
library(caret)
library(e1071)
library(randomForest)
library(gbm)
knitr::opts_chunk$set(echo=TRUE,tidy.opts=list(width.cutoff=70),tidy=TRUE)
```

# Introduction  

## Purpose  

The overall theme of this exercise is determining the popularity of online news. The goal is to create models for predicting the popularity of news articles from *mashable.com* using binary classification to categorize article shares in social networks. Two models will be created: a linear regression model and a non-linear ensemble model. The parameter functionality of markdown will be used to automatically generate an analysis report for each day of the week that an article might be published.  

## Data Description  

The data comes from the *UC Irvine Machine Learning Repository* and can be found in its original form at this [link](https://archive.ics.uci.edu/ml/datasets/Online+News+Popularity#).  

This data set covers a two year period, listing all published articles from Mashable and several variables of associated characteristics, such as:  

*  Published Day of Week
*  Channel, or Genre (Lifestyle, Business, World, Entertainment, etc.)
*  Word Count
*  Number of Images
*  *...and several others...*

In total there are **58 possible predictive attributes** with the goal of predicting the response variable of article popularity using a binary classification based on number of shares in social networks. The provided link above goes into more detail for all variables and their representation.  

## Methods  

The methods used in the creation of both models will follow a similar approach. The cleansed data set for each day of the week will be split into a training (70% of records) and a testing set (30% of records). The training set will undergo k-fold cross validation for the model fit, with a varying level of folds and resampling for each model type. The fitted model will be selected automatically by the `caret` package based on resulting accuracy, and then applied to the testing data set to determine actual accuracy and associated misclassification rate.  

The non-linear ensemble model for this exercise will be Random Forests, and the linear model will be a Logistic Regression model under the family of Generalized Linear Regression. More detail around these two model types is discussed further in their relevant sections.

The final *Conclusion* section will automatically select and display the accuracy and misclassification rate from the best model, determined by accuracy.  

# Automation  

```{r automation, message=FALSE}
# Create a vector of the elements to iterate over
days <- c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")

# Render to .md the template for each param
purrr::map(.x = days,.f=~rmarkdown::render(input="Template.Rmd",params=list(day=.x),output_file=paste0(.x,".md")))
```

A template file has been created that utilizes the above R markdown automation to generate a full analysis with a metadata overview, data import and slicing, numeric summaries, model fit, model selection, model testing, and final conclusion for best model used for each day of the week. These individual analysis can be found within the below links:  

The analysis for [Monday is available here](Monday.md).  
The analysis for [Tuesday is available here](Tuesday.md).  
The analysis for [Wednesday is available here](Wednesday.md).  
The analysis for [Thursday is available here](Thursday.md).  
The analysis for [Friday is available here](Friday.md).  
The analysis for [Saturday is available here](Saturday.md).  
The analysis for [Sunday is available here](Sunday.md).  
